Variants of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus

### 2016-09-16 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Make a few minor improvements

H\'ajek \citep{hajek} claimed that Anderson's emendation makes Axioms 4 and 5 derivable from the other axioms and definitions. Discussions about the redundancy of these two axioms started with an earlier (and wrong claim) claim by Magari \citep{magari} w.r.t. Scott's variant, and the redundancy was the topic of a long controversy \citep{controversy} between H\'ajek and Anderson. As confirmed in Subsection \ref{sec:NonRedundantAnderson}, H\'ajek's claim is correct. Nevertheless, the proof presented above still uses Axioms 4 and 5, since this section's intention is to show that Anderson's emendation requires only a minimal modification of the proofs presented in the previous sections.

@@

In order to construct a natural deduction proof of theorem \ref{A:T3} using only Anderson's axioms \ref{A:A1}, \ref{A:A2} and \ref{A:A3} as well as Anderson's Definition \ref{A:D1}, it suffices to modify 
the proof of Lemma \ref{A:L1}, which is the only part of the argument that depended on axioms \ref{A:A4} and \ref{A:A5} and definitions \ref{A:D2} and \ref{A:D3}. Such a modified proof is shown below.

@@

\begin{proof} \hfill
\begin{prooftree}
\AXC{$ $} \RightLabel{2}
\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{D1}}
\UIC{$\forall \varphi. [P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma)]$}\RightLabel{$\forall_E$}
\UIC{$P(G) \biimp \nec G(\gamma)$} \RightLabel{$\wedge_E$}
\UIC{$P(G) \imp \nec G(\gamma)$}
\AXC{Axiom \ref{A:A3}} \RightLabel{}
\UIC{$P(G)$} \RightLabel{$\imp_E$}
\BIC{$\nec G(\gamma)$}
\AXC{$ $} \RightLabel{1}
\UIC{$G(\gamma)$} \RightLabel{$\ex_I$}
\UIC{$\ex x.G(x)$}  \RightLabel{$\imp_I^1$}
\UIC{$G(\gamma)\imp \ex x.G(x)$} \RightLabel{$\nec_I$}
\UIC{$\nec [G(\gamma)\imp \ex x.G(x)]$} \doubleLine  \RightLabel{(Axiom K)}
\UIC{$\nec G(\gamma)\imp \nec \ex x.G(x)$} \RightLabel{$\imp_E$}
\BIC{$\nec \ex x.G(x)$}\RightLabel{$\imp_I^2$}
\UIC{$\ex z. G(z)\imp \nec \ex x.G(x)$}
\end{prooftree}
%\end{scriptsize}

@@

The proof of section \ref{sec:newproof} shows explicitly that the weak modal system {\bf KB} is sufficient for the ontological argument as opposed to the standard modal calculus {\bf S5}. This weakening of the system is based on an explicit formalization and confirms the result briefly described by Anderson \citep[footnote 2]{and90}. The weakening also applies to the emendation and variant of Anderson and Bj\o rdal respectively. The proofs are short and accessible to the general reader because of the choice of a natural deduction calculus.

>>>>

H\'ajek \citep{hajek} claimed that Anderson's emendation makes Axioms 4 and 5 derivable from the other axioms and definitions. Discussions about the redundancy of these two axioms started with an earlier (and wrong) claim by Magari \citep{magari} w.r.t. Scott's variant, and the redundancy was the topic of a long controversy between H\'ajek and Anderson \citep{controversy}. As confirmed in Subsection \ref{sec:NonRedundantAnderson}, H\'ajek's claim is correct. Nevertheless, the proof presented above still uses Axioms 4 and 5, since this section's intention is to show that Anderson's emendation requires only a minimal modification of the proofs presented in the previous sections.

@@

In order to construct a natural deduction proof of theorem \ref{A:T3} using only Anderson's axioms \ref{A:A1}, \ref{A:A2} and \ref{A:A3} as well as Anderson's Definition \ref{A:D1}, it suffices to modify 
the proof of Lemma \ref{A:L1}, which is the only part of the argument that depended on axioms \ref{A:A4} and \ref{A:A5} and definitions \ref{A:D2} and \ref{A:D3}. The modified proof is shown below.

@@

\begin{proof} \hfill
\begin{prooftree}
\AXC{$ $} \RightLabel{2}
\UIC{$\ex z. G(z)$}\RightLabel{$\ex_E$}
\UIC{$G(\gamma)$}  \dottedLine \RightLabel{D\ref{D1}}
\UIC{$\forall \varphi. (P(\lambda x.\varphi(x)) \biimp \nec \varphi(\gamma))$}\RightLabel{$\forall_E$}
\UIC{$P(G) \biimp \nec G(\gamma)$} \RightLabel{$\wedge_E$}
\UIC{$P(G) \imp \nec G(\gamma)$}
\AXC{Axiom \ref{A:A3}} \RightLabel{}
\UIC{$P(G)$} \RightLabel{$\imp_E$}
\BIC{$\nec G(\gamma)$}
\AXC{$ $} \RightLabel{1}
\UIC{$G(\gamma)$} \RightLabel{$\ex_I$}
\UIC{$\ex x.G(x)$}  \RightLabel{$\imp_I^1$}
\UIC{$G(\gamma)\imp \ex x.G(x)$} \RightLabel{$\nec_I$}
\UIC{$\nec (G(\gamma)\imp \ex x.G(x))$} \doubleLine  \RightLabel{(Axiom K)}
\UIC{$\nec G(\gamma)\imp \nec \ex x.G(x)$} \RightLabel{$\imp_E$}
\BIC{$\nec \ex x.G(x)$}\RightLabel{$\imp_I^2$}
\UIC{$\ex z. G(z)\imp \nec \ex x.G(x)$}
\end{prooftree}
%\end{scriptsize}

@@

The formal proofs presented here are short and accessible because of the choice of a natural deduction calculus. The new proof of section \ref{sec:newproof} shows explicitly that the weaker modal logic {\bf KB} is sufficient for the ontological argument. The stronger modal logic {\bf S5}, which is usually taken for granted in discussions of G\"odel's ontological argument, is not necessary. This weakening of the underlying logic confirms a claim briefly mentioned by Anderson \citep[footnote 2]{and90}. {\bf KB} is also sufficient for the variants of Anderson and Bj\o rdal, which address Sobel's criticism of the modal collapse.
